# Reproducible Research: Peer Assessment 1

This is the peer Assessment 1 for the coursera course "Reporduceable Research"
at [coursera](https://class.coursera.org/repdata-002). The detailed instructions are available
at the [Coursera Peer Assessment Page](https://class.coursera.org/repdata-002/human_grading/view/courses/972084/assessments/3/submissions).  
The Work is based on a fork from a provided [Github repository](http://github.com/rdpeng/RepData_PeerAssessment1). All nessesary data and template files are included in the Github repository.   
Please read the [README.md](README.md) file for further information.

## System Setup
```{r results='hide'}
options(digits = 4, scipen = 4);
opts_chunk$set(echo = TRUE);
require("utils");
require("graphics");
require("grDevices");
require("lattice");
rm(list=ls());
#set timeformat to english on windows systems.
Sys.setlocale(category = "LC_TIME", locale = "english");
```

## Loading and preprocessing the data

```{r}
Amd <- read.csv(
    file=unzip(zipfile="activity.zip",files=c("activity.csv"))
    ,header=TRUE
    ,sep=","
    ,quote="\""
    ,dec="."
    ,na.strings= "NA"
    ,stringsAsFactors = TRUE);
colnames(Amd)<-c("steps","datefactors","interval");
Amd$steps<-as.numeric(Amd$steps);
Amd$date<-strptime(Amd$date,format="%Y-%m-%d");
AmdTidy <- Amd[!is.na(Amd$steps), ];
```


```{r HistogramNARemoved}
AmdbyDate<-aggregate.data.frame(
    x=AmdTidy[,"steps"]
    ,by=list(datefactors=AmdTidy$datefactors)
    ,FUN=sum);
colnames(AmdbyDate)[2]<-"steps";

#https://class.coursera.org/repdata-002/forum/thread?thread_id=49
#discussion about hist or barplot - i started with barplot
hist(x=AmdbyDate$steps,col="blue",xlab="Steps",main="Histogram of steps");
```

## What is mean total number of steps taken per day?

```{r}
MeanOfStepsPerDay<-round(mean(x=AmdbyDate$steps),4);
MedianOfStepsPerDay<-round(median(x=AmdbyDate$steps),4);

```
## What is the average daily activity pattern?

```{r LineChartNARemoved}
AmdbyInterval<-aggregate.data.frame(
    x=AmdTidy[,"steps"]
    ,by=list(interval=AmdTidy$interval)
    ,FUN=mean);
colnames(AmdbyInterval)[2]<-"meansteps";

plot(
    x=AmdbyInterval$interval
    ,y=AmdbyInterval$meansteps
    ,type="l"
    ,xlab="interval"
    ,ylab="Avg steps"
    ,main="Average steps by interval");

MaxIntervalValue <- max(AmdbyInterval$meansteps);
MaxIntervalNumber <- AmdbyInterval[AmdbyInterval$meansteps==MaxIntervalValue,"interval"];
```
## Imputing missing values

```{r HistogramNASubstituted}
#NumberOfNaValues <- sum(!complete.cases(Amd[,c("steps","interval","datefactors")]));
NumberOfNaValues <- sum(is.na(Amd$steps));

AmdFull<-(merge(x=Amd,y=AmdbyInterval,by="interval"));
#For easier error checking sort the dataframe
AmdFull<-AmdFull[order(AmdFull$date,AmdFull$interval),];
NAVector<-is.na(AmdFull$steps);
AmdFull[NAVector,"steps"]<-AmdFull[NAVector,"meansteps"];
AmdFull$meansteps<-NULL;

AmdFullbyDate<-aggregate.data.frame(
    x=AmdFull[,"steps"]
    ,by=list(datefactors=AmdFull$datefactors)
    ,FUN=sum);
colnames(AmdFullbyDate)[2]<-"steps";

hist(x=AmdFullbyDate$steps,col="blue",xlab="Steps",main="Histogram of steps");

MeanOfStepsPerDayFull<-round(mean(x=AmdFullbyDate$steps),4);
MedianOfStepsPerDayFull<-round(median(x=AmdFullbyDate$steps),4);

```
## Are there differences in activity patterns between weekdays and weekends?

```{r XYPlotNASubstituted}

AmdFull$Weekday<-weekdays(x=AmdFull$date,abbreviate=TRUE);
AmdFull[AmdFull$Weekday=="Sun" | AmdFull$Weekday=="Sat","Weekday"]<-"Weekend";
AmdFull[AmdFull$Weekday!="Weekend","Weekday"]<-"Weekday";
AmdFull$Weekday<-as.factor(AmdFull$Weekday);

AmdPanelPlot<-aggregate.data.frame(
    x=AmdFull$steps
    ,by=list(Weekday=AmdFull$Weekday,interval=AmdFull$interval)
    ,FUN=mean);
colnames(AmdPanelPlot)[3]<-"steps";

xyplot(steps ~ interval | Weekday,
       data = AmdPanelPlot,
       type = 'l',
       layout = c(1,2), 
       col = c('blue'),
       main = "Avg steps by time", 
       xlab= "Interval", 
       ylab = "Number of steps"
);
```